Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-03 Paper/paper.tex
pass on intro

\graphicspath{{Figures/}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}

@@

Repetitive, periodic motions are ubiquitous in our world, such as animal locomotion (walking/flapping/slithering/swimming), mechanical motions (wheels spinning, pistons oscillating), biological rhythms (heartbeat, breathing), musical rhythms (drumming, feet tapping), etc.  In this work, we consider the problem of synthesizing a fine detail, ``slow motion'' template cycle of motion from a video of such a motion with many periods.  This fine detailed analysis can be applied, for example, to characterize subtle progressions of blood flow in the face during a heartbeat \cite{kumar2015distanceppg}.  Having access to a consensus template can also be used to visualize variations from cycle to cycle, such as in a repetitive automated action on an assembly line, where too much variation could indicate the onset of failure, in repetitive human motions to assess the stress of those motions \cite{greene2017visualizing}, and in athletic activities for optimizing performance.

Our approach to slow motion templates is {\em Eulerian}; that is, we process the video pixel by pixel with no tracking.  Eulerian approaches for video synthesis are attractive due to their simplicity and ease of implementation.  For instance, a simple Fourier bandpass filtering and amplifying pixels can successfully elucidate subtle periodic motions in videos \cite{wu2012eulerian, wadhwa2013phase}.  Additionally, time domain Eulerian approaches have been used to synthesize infinitely playing ``video textures'' using random walks \cite{schodl2000video}, or to synthesize perfectly looping video templates after grouping pixels with similar periods together and devising a common period for all groups \cite{Liao2013VideoLoops,Liao2015VideoLoops}.  There are some pitfalls with Eulerian techniques, however.  Purely Eulerian amplification techniques are known to fail for large motions, and they also require the user to manually specify frequencies of interest \cite{wu2012eulerian, wadhwa2013phase}.  Furthermore, drift from period to period due to shaking and varying motions from cycle to cycle can present a problem, as can unrelated motion in the background \cite{stauffer1999adaptive}.

To address the problems with an Eulerian representation, we devise a geometric/topological Eulerian framework byconstructing {\em sliding window embeddings} of our videos (Section~\ref{sec:slidingwindow}).  Sliding window embeddings, or ``delay reconstructions,'' have found a diverse array of applications in activity recognition \cite{frank2010activity,venkataraman2016shape}, gene expression data \cite{perea2015sw1pers}, EEG analysis \cite{stam2005nonlinear, plesnik2014detection}, audio and music analysis \cite{herzel1994analysis,serra2009cross,bello2011measuring,traliemoebius}, video analysis \cite{tralie2017quasi}, and motion capture analysis \cite{venkataraman2016shape}.  Sliding window embeddings of periodic time series, in particular, form samples of a topological loop \cite{perea2015sliding}, as do generalized multivariate sliding windows on periodic video data, regardless of the type of motion present \cite{traliehigh, tralie2017quasi}.  Furthermore, as shown by the authors of \cite{tralie2017quasi}, sliding window embeddings on Eulerian videos provide a sort of ``time regularization,'' which can mitigate the effects of drift, leading to a cleaner loop.  Our geometric framework also allows us to use topological data analysis \cite{edelsbrunner2000topological,edelsbrunner2008persistent,edelsbrunner2010computational,carlsson2009topology,ghrist2014elementary} and fundamental frequency estimation \cite{Mcleod05asmarter} to autotune the spatial and temporal scales of the periodic motion, respectively, so that no user intervention is required.  Our embedding also makes it easy to assign each window a circular phase $\phi \in [0, 2\pi]$ using Laplacian Eigenmaps \cite{belkin2003laplacian}.  We then re-order the sliding windows by $\phi$ and exploit the sliding window structure to have different windows vote on the final pixels in each frame of the template, which further mitigates drift (Section~\ref{sec:cyclereordering}).  As an illustration, Figure~\ref{fig:ConceptFigure} shows an example of our technique on a sampled periodic time series (i.e. single pixel grayscale video).  There are only 12 samples per period in the original signal, so the details of each period coarse and noisy.  However, once they are re-sorted, we get a nice, fine-detailed representation of one period.  The extension of this technique to videos will be described in detail in Section~\ref{sec:methods}.

Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, and quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result.  Our use of the graph Laplcian to parameterize time series is also unique, though we are inspired by works that have used it to re-arrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and works which use vector diffusion maps to re-order the frames of microscope images a developing embryo around a loop \cite{dsilva2015diffusionvecordering}.  One can also view the combination of TDA + Laplacian eigenmaps as a simpler alternative to cohomology circular coordinates \cite{de2011persistent,vejdemo2015cohomological} for the special case of periodic data.  It is also the opposite of the approach in \cite{bendich2011improving}, which performs diffusion maps before TDA.

>>>>

\usepackage{tikz}

\graphicspath{{Figures/}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\matt}[1]{{\color{purple}[MSB: #1]}}

@@

Repetitive, periodic behavior is ubiquitous in our world. Animal locomotion, mechanical motion, biological rhythms, and musical rhythms can all be characterized as periodic phenomena.  In this work, we consider the problem of taking a video that contains such repetitive motion of many periods, and synthesizing a single fine detail, slow motion template cycle.  This fine detailed analysis can be applied, for example, to characterize subtle progressions of blood flow in the face during a heartbeat \cite{kumar2015distanceppg}.  Having access to a consensus template can also be used to visualize variations from cycle to cycle. This could be used to indicate the onset of failure in a repetitive automated action on an assembly line, to assess the stress of motions in repetitive human actions~\cite{greene2017visualizing}, and to optimize performance in athletic activities.

Our approach to slow motion templates is {\em Eulerian}; that is, we process the video pixel by pixel with no tracking.  Eulerian approaches for video synthesis are attractive due to their simplicity and ease of implementation.  For instance, a simple Fourier bandpass filtering and amplifying pixels can successfully elucidate subtle periodic motions in videos \cite{wu2012eulerian, wadhwa2013phase}.  Additionally, time domain Eulerian approaches have been used to synthesize infinitely playing ``video textures'' using random walks \cite{schodl2000video}, or to synthesize perfectly looping video templates after grouping pixels with similar periods together and devising a common period for all groups \cite{Liao2013VideoLoops,Liao2015VideoLoops}.  There are some pitfalls with Eulerian techniques, however.  They are known to fail for large motions, and also require the user to manually specify frequencies of interest \cite{wu2012eulerian, wadhwa2013phase}.  Furthermore, period-to-period drift that is due to a moving camera or small variation in motion can be problematic, as well as unrelated motion in the background \cite{stauffer1999adaptive}.

% MSB: opportunity to slim paper by cutting examples 
To address the problems with an Eulerian representation, we devise a geometric/topological Eulerian framework by constructing {\em sliding window embeddings} of videos (Section~\ref{sec:slidingwindow}).  Sliding window embeddings, or delay reconstructions, have found a diverse array of applications in activity recognition \cite{frank2010activity,venkataraman2016shape}, gene expression data \cite{perea2015sw1pers}, EEG analysis \cite{stam2005nonlinear, plesnik2014detection}, audio and music analysis \cite{herzel1994analysis,serra2009cross,bello2011measuring,traliemoebius}, video analysis \cite{tralie2017quasi}, and motion capture analysis \cite{venkataraman2016shape}.  Sliding window embeddings of periodic time series, in particular, form samples of a topological loop \cite{perea2015sliding}, as do generalized multivariate sliding windows on periodic video data, regardless of the type of motion present \cite{traliehigh, tralie2017quasi}.  Furthermore, the sliding window embedding of an Eulerian video provides a form of regularization that mitigates the effects of drift and leads to a cleaner loop~\cite{tralie2017quasi}.  Our geometric framework also allows us to use topological data analysis \cite{edelsbrunner2000topological,edelsbrunner2008persistent,edelsbrunner2010computational,carlsson2009topology,ghrist2014elementary} and fundamental frequency estimation \cite{Mcleod05asmarter} to autotune the spatial and temporal scales of the periodic motion, respectively, so that no user intervention is required.  We use Laplacian Eigenmaps~\cite{belkin2003laplacian} computed on the sliding window embedding to obtain a circular phase $\phi \in [0, 2\pi]$ for each point in the embedding.  The video is re-ordered by using $\phi$ and exploiting the sliding window structure, where different windows vote on the final pixels in each frame of the template, helping to further mitigate drift (Section~\ref{sec:cyclereordering}).  As an illustration, Figure~\ref{fig:ConceptFigure} shows an example of our technique on a sampled periodic time series (i.e. single pixel grayscale video).  There are only 12 samples per period in the original signal, so the details of each period are coarse and noisy.  However, once they are re-sorted, we get a fine-detailed representation of one period.  The extension of this technique to videos will be described in detail in Section~\ref{sec:methods}.

% MSB: opportunity to slim paper by cutting examples 
Many prior works use a combination of sliding window embeddings and topological data analysis, such as detecting the Lorenz attractor \cite{de2012topological}, quantifying circadian rhythms in gene expression data \cite{perea2015sliding}, detecting chatter in mechanical systems \cite{khasawneh2016chatter}, detecting wheezing in audio \cite{emrani2014real}, quantifying periodic activites in motion capture data \cite{vejdemo2015cohomological, venkataraman2016persistent}, and quantifying vocal fold anomalies in video \cite{tralie2017quasi}, though we believe we are the first to use a sliding window + TDA framework to {\em synthesize} a new result. Related works have used the graph Laplacian to rearrange images around a loop as a pre-processing step for structure from motion \cite{averbuch2015ringit}, and to reorder the frames of microscope images from a developing embryo around a loop using vector diffusion maps~\cite{dsilva2015diffusionvecordering}.  However, our use of the graph Laplacian to instead {\em parameterize} time series is unique. One can also view the combination of TDA + Laplacian eigenmaps as a simpler alternative to cohomology circular coordinates \cite{de2011persistent,vejdemo2015cohomological} for the special case of periodic data.  It is also the opposite of the approach in \cite{bendich2011improving}, which performs diffusion maps before TDA. \matt{unclear what ''It'' refers to in this last sentence}
